State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-16 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
edits from David up to results

In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements, possibly preventing us from positioning sensors in accordance to equation~\ref{eq:MinimumSensorDistance}. However, we envisage it may be possible to estimate it using other modalities with higher spatial resolution such as, fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Nevertheless, spectral aliasing can still be avoided by a proper choice of the spatial sampling distance given the sensors spectral characteristics. The spatial extent of the sensors results in spectral low-pass action, thus providing spatial anti-aliasing filtering. Such sensors therefore can only sense the field up to a known bandwidth, denoted by $\nu_{cy}$. Aliasing can then be avoided by positioning the sensors according to equation~\ref{eq:MinimumSensorDistance}, with $\nu_{cy}$ replacing $\nu_c$.

Although such sensors avoid errors due to aliasing, they attenuate the high spatial frequency variations in the observations. Therefore, any procedure applied to estimate the original field or the underlying connectivity structure from these band-limited observations has the potential to underestimate the high spatial frequency components of both the field and the connectivity kernel. Such a remark motivates the use of sensors with wider bandwidths (narrower in space). Nevertheless, this choice would require more sensors in order to satisfy Shannon's sampling theorem for a reasonable spatial region. Therefore, a compromise needs to be found between the bandwidth of the sensor response, the accuracy of the estimation results, the number of sensors used and the computational demands of the estimation procedure when designing experiments.

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. Again applying Shannon's Sampling Theorem, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} < \frac{1}{2\rho_{\phi}\nu_{cy}}
\end{equation}
where $\rho_{\phi}$ is an oversampling parameter to determine the basis function separation.

The width parameter of the Gaussian basis functions can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. To demonstrate this we begin with the 2-dimensional Gaussian basis function defined as
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
with the corresponding Fourier transform
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{cy}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

Note that for a spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Thus, the knowledge of the distance between basis functions directly implies the total number of basis functions required to represent a known spatial region.

@@

The unscented Rauch Tung Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{equation}\label{eq:sigmapoints1}
	\mathcal X_{0}=\bar x 
\end{equation}
\begin{equation}
	\mathcal X_{i}=\bar x+(\sqrt{( L + \lambda)\mathbf P_x})_i, \quad i=1, \dots, L 
\end{equation}
\begin{equation}\label{eq:sigmapoints2}
	\mathcal X_{i}=\bar x-(\sqrt{( L + \lambda)\mathbf P_x})_{i- L}, \quad i= L+1, \dots, 2 L 
\end{equation}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix from the filtering or smoothing, $(\sqrt{( L + \lambda)\mathbf P_x})_i$ is the $i^{th}$ column of the weighted matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points, $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$ and $\kappa$ is a scaling parameter (see~\cite{Haykin2001} for more details).

@@

	\mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L. 

@@

where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the cross-covariance matrix of the states, $\mathbf M$, needs to be computed. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that  the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used, otherwise the state vector should be augmented with the noise terms. The additive disturbance form reduces the dimension and the number of sigma points used in the algorithm.

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new state vector sequence estimate. To create the least squares parameter estimator we first define the $L \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from the URTSS, we can write
\begin{eqnarray*}
	\mathbf x_{1} &=& \mathbf{q}(\mathbf x_0) \boldsymbol{\theta}+\xi\mathbf x_0+\mathbf e_0 \\
	\mathbf x_{2} &=& \mathbf{q}(\mathbf x_1) \boldsymbol{\theta}+\xi\mathbf x_1+\mathbf e_1  \\
	&\vdots& \\
	\mathbf x_{T}&=&\mathbf{q}(\mathbf x_{T-1}) \boldsymbol{\theta}+\xi\mathbf x_{T-1}+\mathbf e_{T-1}. 
\end{eqnarray*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathbf W+\mathbf{e} 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\mathbf x_{1}\\
		\mathbf x_{2}\\
		\vdots\\
		\mathbf x_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\mathbf x_0)& \mathbf x_{0}\\
		\mathbf q(\mathbf x_1)& \mathbf x_{1}\\
		\vdots & \vdots\\
		\mathbf q(\mathbf x_{T-1})& \mathbf x_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathbf W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the parameters, $ \mathbf W$, can be estimated using
\begin{equation}
	\mathbf W\approx(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}
% The accuracy of the approximation is linked to the time step used in the temporal discretisation of the system, where the state transition appears more linear with a finer resolution. 
% Alternatives to the UKF are the Extended Kalman Filter (EKF)~\cite{Haykin2001} and the Sequential Monte Carlo (SMC) filter~\cite{doucet2001}. The EKF approximates the state transition equation by linearising about the current state estimates (by calculating the Jacobian) at each time instance. The linearisation maintains the Gaussianality of the model. The unscented Kalman filter approximates the posterior state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, while maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, since the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the EKF can be problematic. SMC filtering can theoretically provide an exact posterior state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to high computation demands.

%The method is appropriate for state estimation of stochastic nonlinear dynamical systems. 
\section{Results}\label{ResultsSection} The full neural field model derived in Section~\ref{NeuralModelSection} was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Model Parameters}. Note, the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is due to the fact that the temporal aspects scale with the sampling rate and the length of the times series and the spatial aspects scale with the size of the neural field and the space step used for numerical simulation. Figure~\ref{fig:experimental design}(a) shows example data from the sensors of the model, and Figure~\ref{fig:experimental design}(b) shows the power spectral density (PSD) of the simulated time-series. The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.
\begin{table}\footnotesize
\begin{tabular}{ccl}
	\hline\hline Symbol & Value & Units \\
	\hline\hline
	$\tau$ & 0.01 & s$^{-1}$\\
	$f_{max}$ & 10 & spikes/s \\
	$\varsigma$ & 0.8 & spike/mV\\
	$v_0$ & 2 & mV\\
	$\boldsymbol{\theta}$ & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\right]^{\top}$ & mV/spike\\
	$\sigma_{\psi_{1}}, \sigma_{\psi_{2}}, \sigma_{\psi_{3}}$ & 1.8, 2.4, 6 & mm\\
	$\sigma_{m},\Delta_{y}$&0.9, 1.5&mm\\
	$\Sigma_{\epsilon}$ &$0.1 \times I_{N} $& mm$^2$ \\
	$\sigma_{\gamma}$&1.3&mm\\
	$\Delta$ & 0.5 & mm \\
	$T_s$ & 1 & ms \\ 
\end{tabular}
\caption{Parameters for the neural field model, which are used to generate synthetic data.}
\label{tab:Model Parameters}
\end{table}
The sensors were spaced in a regular square 14 $\times$ 14 grid, the sensor width $\sigma^2_m$ was 0.81 which equals $1.5$~mm full width at half maximum, with spacing $\Delta_y$ equal to $1.5$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. 

@@

   	\caption{\textbf{a.} Example of data generated by the full neural field model from the first 5 sensors. \textbf{b.} Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of iEEG.} 

@@

\caption{Spatial frequency of the neural fields. \textbf{a.} The average (over time) spatial frequency of the neural field. \textbf{b.} The average (over time) spatial frequency of the estimated neural field. As a result of the band limiting properties of the sensors and basis functions the reconstructed field shows a narrower bandwidth.}

@@

The spatial frequency from the observations is shown in Figure~\ref{fig:ObservationFreqAnal}. The cut-off frequency, $\nu_{cy}$, was taken to be $0.19$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:SensorFunctionFreqResponse}. Using an oversampling parameter, $\rho_{\phi}$, of $1$, the minimum distance between basis functions was calculated to be $2.63$ to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}).\parham{In order to achieve a smooth representation of the field and alleviate the effect of the disturbance we set the cut-off frequency of the basis functions lower than the observation cut-off frequency to 0.12. It should be noted that the smooth characteristic of Gaussian low pass filter allows  passing frequencies higher than the cut-off frequency.  From equation~\ref{eq:WidthCutOffRelationship} the required basis function width to represent the dominant spatial frequencies was calculated to be $2.5$.The distance between basis functions, $\Delta_{\phi}$, was set to $2.5$, which is sufficient to capture the significant dynamics according to our criteria.} From equation~\ref{eq:WidthCutOffRelationship} the required basis function width to represent the dominant spatial frequencies was calculated to be $1$. The distance between basis functions, $\Delta_{\phi}$, was set to $2.5$, which is sufficient to capture the significant dynamics according to our criteria. The basis function width, $\sigma_{\phi}^2$, was set above the calculated width to $2.5$, which reflected a design choice. This was a result of a trade-off that ensured that the lower spatial frequency components of the field were better represented at the expense of higher spatial frequency components \mike{I don't know why there's a tradeoff between estimating low and high frequency components of the kernel, and I'm not sure I've been convinced by this section}. The spatial frequency response of the basis function can be seen in Figure~\ref{fig:BasisFunctionFreqResponse}.

>>>>

In practice, it is difficult to estimate the bandwidth of the cortex using traditional electrophysiological measurements, possibly preventing us from positioning sensors according to equation~\ref{eq:MinimumSensorDistance}. However, we envisage it may be possible to estimate the spatial bandwidth using other modalities with higher spatial resolution, such as fMRI, NIRS or other optical imaging techniques~\cite{Issa2000}. Nevertheless, spectral aliasing can still be avoided by a proper choice of the spatial sampling distance given the sensors' spectral characteristics. The spatial extent of the sensors results in spectral low-pass action, thus providing spatial anti-aliasing filtering. Such sensors therefore can only sense the field up to a known bandwidth, denoted by $\nu_{cy}$. Aliasing can then be avoided by positioning the sensors according to equation~\ref{eq:MinimumSensorDistance}, with $\nu_{cy}$ replacing $\nu_c$.

Although such sensors avoid errors due to aliasing, they attenuate the high spatial frequency variations in the observations. Therefore, any procedure applied to estimate the original field or the underlying connectivity structure from these band-limited observations has the potential to underestimate the high spatial frequency components of both the field and the connectivity kernel. This motivates the use of sensors with wider bandwidths (narrower in space). Nevertheless, this choice would require more sensors in order to satisfy Shannon's sampling theorem for a reasonable spatial region. Therefore, a compromise needs to be found between the bandwidth of the sensor response, the accuracy of the estimation results, the number of sensors used and the computational demands of the estimation procedure when designing experiments.

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. Again applying Shannon's Sampling Theorem, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} < \frac{1}{2\rho_{\phi}\nu_{cy}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

The width parameter of the Gaussian basis functions can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. To demonstrate this, we begin with the two-dimensional Gaussian basis function defined as
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
with the corresponding Fourier transform
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{cy}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent the field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

Note that for spatially homogeneous isotropic fields, the basis functions can be placed on a regular grid. Thus, the knowledge of the distance between basis functions can be used to determine the total number of basis functions required to represent a known spatial region.

@@

The unscented Rauch Tung Striebel smoother (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward iteration to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward iteration to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and the second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{equation}\label{eq:sigmapoints1}
	\mathcal X_{0}=\bar x 
\end{equation}
\begin{equation}
	\mathcal X_{i}=\bar x+\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, L 
\end{equation}
\begin{equation}\label{eq:sigmapoints2}
	\mathcal X_{i}=\bar x-\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_{i- L}, \quad i= L+1, \dots, 2 L 
\end{equation}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix from the filtering or smoothing, $\left(\sqrt{( L + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the weighted matrix square root of $\mathbf P_x$ and $L$ is the dimension of the state space. The total number of sigma points is $2L+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( L+\kappa) - L, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points, $\beta$ incorporates prior knowledge of the distribution of $\mathbf{x}$ and $\kappa$ is a scaling parameter (see~\cite{Haykin2001} for more details).

@@

	\mathbf W_i^{(m)}=\mathbf W_i^{(c)}=\frac{1}{2( L+\lambda)} \quad i=1, \dots, 2L, 

@@

where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the cross-covariance matrix of the states, $\mathbf M$, needs to be computed. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that the disturbance, $\mathbf{e}_t$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTSS is used, otherwise the state vector should be augmented with the noise terms. The additive disturbance form reduces the dimension and the number of sigma points used in the algorithm.

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new state vector sequence estimate. To create the least squares parameter estimator, we first define the $L \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from the URTSS, we can write
\begin{eqnarray*}
	\mathbf x_{1} &=& \mathbf{q}(\mathbf x_0) \boldsymbol{\theta}+\xi\mathbf x_0+\mathbf e_0 \\
	\mathbf x_{2} &=& \mathbf{q}(\mathbf x_1) \boldsymbol{\theta}+\xi\mathbf x_1+\mathbf e_1  \\
	&\vdots& \\
	\mathbf x_{T}&=&\mathbf{q}(\mathbf x_{T-1}) \boldsymbol{\theta}+\xi\mathbf x_{T-1}+\mathbf e_{T-1}. 
\end{eqnarray*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathbf W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\mathbf x_{1}\\
		\mathbf x_{2}\\
		\vdots\\
		\mathbf x_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\mathbf x_0)& \mathbf x_{0}\\
		\mathbf q(\mathbf x_1)& \mathbf x_{1}\\
		\vdots & \vdots\\
		\mathbf q(\mathbf x_{T-1})& \mathbf x_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathbf W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the parameter estimates, $ \mathbf{\hat{W}}$, are
\begin{equation}
	\mathbf{\hat{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}
% The accuracy of the approximation is linked to the time step used in the temporal discretisation of the system, where the state transition appears more linear with a finer resolution. 
% Alternatives to the UKF are the Extended Kalman Filter (EKF)~\cite{Haykin2001} and the Sequential Monte Carlo (SMC) filter~\cite{doucet2001}. The EKF approximates the state transition equation by linearising about the current state estimates (by calculating the Jacobian) at each time instance. The linearisation maintains the Gaussianality of the model. The unscented Kalman filter approximates the posterior state density by a Gaussian distribution using a minimal set of carefully chosen sigma points, while maintaining the nonlinearity in the system. This has been shown to give superior performance over the EKF in state estimation, since the EKF maintains a first order approximation where the UKF provides an approximation accurate at least to the second order. In addition, the computation of the Jacobian used in the EKF can be problematic. SMC filtering can theoretically provide an exact posterior state density for nonlinear systems. However, currently this method is not appropriate for our problem, due to high computation demands.

%The method is appropriate for state estimation of stochastic nonlinear dynamical systems. 
\section{Results}\label{ResultsSection} The full neural field model derived in Section~\ref{NeuralModelSection} was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Model Parameters}. Note, the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is due to the fact that the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the space step used for numerical simulation. Figure~\ref{fig:experimental design}(a) shows example data from the sensors of the model and Figure~\ref{fig:experimental design}(b) shows the power spectral density (PSD) of the simulated time-series. The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.
\begin{table}\footnotesize
\centering
\begin{tabular}{ccl}
	\hline\hline Symbol & Value & Units \\
	\hline\hline
	$\tau$ & 0.01 & s$^{-1}$\\
	$f_{max}$ & 10 & spikes/s \\
	$\varsigma$ & 0.8 & spike/mV\\
	$v_0$ & 2 & mV\\
	$\boldsymbol{\theta}$ & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\end{array}
	\right]^{\top}$ & mV/spike\\
	$\sigma_{\psi_{1}}, \sigma_{\psi_{2}}, \sigma_{\psi_{3}}$ & 1.8, 2.4, 6 & mm\\
	$\sigma_{m},\Delta_{y}$&0.9, 1.5&mm\\
	$\Sigma_{\epsilon}$ &$0.1 \times I_{N} $& mm$^2$ \\
	$\sigma_{\gamma}$&1.3&mm\\
	$\Delta$ & 0.5 & mm \\
	$T_s$ & 1 & ms \\ 
\end{tabular}
\caption{Parameters for the neural field model that are used to generate synthetic data.}
\label{tab:Model Parameters}
\end{table}

The sensors were spaced in a regular square 14 $\times$ 14 grid, the sensor width $\sigma^2_m$ was $0.81$, which equals $1.5$~mm full width at half maximum, with spacing $\Delta_y$ equal to $1.5$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. 

@@

   	\caption{\textbf{a.} Example of data generated by the full neural field model from the first 5 sensors. \textbf{b.} Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 

@@

\caption{Power in dB as a function of spatial frequency of the neural field. \textbf{a.} The average (over time) power of the spatial frequency of the neural field. \textbf{b.} The average (over time) power of the spatial frequency of the reconstructed neural field from the basis function decomposition. The reconstructed field shows a narrower bandwidth as a result of the band limiting properties of the sensors and basis functions.}

@@

The spatial frequency from the observations is shown in Figure~\ref{fig:ObservationFreqAnal}. The cut-off frequency, $\nu_{cy}$, was taken to be $0.19$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:SensorFunctionFreqResponse}. Using an oversampling parameter, $\rho_{\phi}=1$, the minimum distance between basis functions was calculated to be $2.63$ to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}).\parham{In order to achieve a smooth representation of the field and alleviate the effect of the disturbance we set the cut-off frequency of the basis functions lower than the observation cut-off frequency to 0.12. It should be noted that the smooth characteristic of Gaussian low pass filter allows  passing frequencies higher than the cut-off frequency.  From equation~\ref{eq:WidthCutOffRelationship} the required basis function width to represent the dominant spatial frequencies was calculated to be $2.5$.The distance between basis functions, $\Delta_{\phi}$, was set to $2.5$, which is sufficient to capture the significant dynamics according to our criteria.} From equation~\ref{eq:WidthCutOffRelationship} the required basis function width to represent the dominant spatial frequencies was calculated to be $1$. The distance between basis functions, $\Delta_{\phi}$, was set to $2.5$, which is sufficient to capture the significant dynamics according to our criteria. The basis function width, $\sigma_{\phi}^2$, was set above the calculated width to $2.5$, which reflected a design choice. This was a result of a trade-off that ensured that the lower spatial frequency components of the field were better represented at the expense of higher spatial frequency components \mike{I don't know why there's a tradeoff between estimating low and high frequency components of the kernel, and I'm not sure I've been convinced by this section}. The spatial frequency response of the basis function can be seen in Figure~\ref{fig:BasisFunctionFreqResponse}.
